---
title: "Diary Entry"
author: "Lew Bo Cong"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Diary Entry (Week 11)

### (1) List the visualizations that you are going to use in your project (Answer: What are the variables that you are going to plot? How will it answer your larger question?)

#### **For multiple hospitals**

-   Line charts
    -   plot Dates against Waiting time, Bed Occupancy Rate, and EMD
        Attendance number, in varying weeks or months.
-   Box plots (Horizontal)
    -   plot Hospitals against Waiting time, Bed Occupancy Rate, and EMD
        Attendance number, in varying weeks, months, or years.
-   Bar charts (Horizontal)
    -   plot Hospitals against Bed Occupancy Rate, and EMD Attendance
        number, on different days.

#### **For individual hospital**

-   Calendar heatmap using facet wrap
    -   plot Waiting time arranged by Days and Week number.
-   Line chart overlaying bar chart
    -   plot Dates against Waiting time (line) and Bed Occupancy Rate
        (bar), in varying weeks or months.
    -   plot Dates against Waiting time (line) and EMD Attendance number
        (bar), in varying weeks or months.
-   Scatterplot
    -   plot Waiting time against Bed Occupancy Rate, in varying weeks,
        months or year.
    -   plot Waiting time against EMD Attendance number, in varying
        weeks, months or year.

By utilising the above-mentioned data visualisations, we can look for
possible correlations and patterns that explain the long waiting times
for admission to a ward.

### (2) How do you plan to make it interactive? (Answer: features of ggplot2/shiny/markdown do you plan to use to make the story interactive)

I plan to use shiny to create an interactive app that allows the user to
input a date range and select hospitals to be drawn on the
above-mentioned plots. I wish to use ggplot2 to display relevant
information on the screen while the user hovers over the data points.

### (3) What concepts incorporated in your project were taught in the course and which ones were self-learnt? (Answer: Create a table with topics in one column and Weeks in the other to indicate which concept taught in which week is being used. Leave the entry of the Week column empty for self-learnt concepts)

+----------------------------------------------------+---------------+
| Topics                                             | Weeks         |
+====================================================+:=============:+
| -   Data variables and their types: transforming   | 3             |
|     data into correct type for manipulation        |               |
|                                                    |               |
|     -   e.g. using as.numeric()                    |               |
|<br>                                                |               |
+====================================================+:=============:+
| -   Using variable type `factor` and `date` to     |               |
|     help order categorical variables               |               |
|                                                    |               |
|     -   e.g. fct_inorder( )                        |               |
|                                                    |               |
|     -   fct_reorder()                              |               |
|                                                    |               |
|     -   parse_date_time()                          |               |
|                                                    |               |
|     -   %U                                         |               |
|<br>                                                |               |
+====================================================+:=============:+
| -   Manipulating Data: understood what is exactly  | 4, 9          |
|     a tidy dataset, and the operations to tidy,    |               |
|     manipulate and reshape data, especially via    |               |
|     pipe operator for cleaner code                 |               |
|                                                    |               |
|     -   e.g. mutate()                              |               |
|<br>                                                |               |
+====================================================+:=============:+
| -   Using other functions to tidy data             |               |
|                                                    |               |
|     -   e.g. gather()                              |               |
|                                                    |               |
|     -   relocate()                                 |               |
|                                                    |               |
|     -   cbind()                                    |               |
|<br>                                                |               |
+====================================================+:=============:+
| -   Data visualisation via ggplot2: applying the   | 7             |
|     different ways to plot data for visualisation  |               |
|<br>                                                |               |
+====================================================+:=============:+
| -   Learnt and used additional methods of          |               |
|     visualising data                               |               |
|                                                    |               |
|     -   e.g. various visualisations                |               |
|                                                    |               |
|     -   pie chart                                  |               |
|                                                    |               |
|     -   overlay line chart over boxplot            |               |
|<br>                                                |               |
+====================================================+:=============:+
| -   Learnt essential function arguments to         |               |
|     visualise data in the desired way              |               |
|                                                    |               |
|     -   eg. `aes(group = 1)` as `geom_line()`      |               |
|         argument to draw line between data points  |               |
|                                                    |               |
|     -   remove legend                              |               |
|                                                    |               |
|     -   adding labels to data points and charts    |               |
|                                                    |               |
|     -   `ggplotly` customisation via changing      |               |
|          hover text and preventing subtitle from   |               |
|          disappearing                              |               |
|<br>                                                |               |
+====================================================+:=============:+

### (4) Challenges and errors that I faced and how I overcame them.

The main challenges and errors were trying to present data in a logical
and meaningful way, by customising conventional data plots in the way I
want. Whenever I faced the issue of needing to add or remove features in
my plots, I will search for the solution on Google and experiment it in
my code.
<br>
I would like to show the bed capacity of each public hospital, perhaps
on a map, but this data is only available reliably and accurately in the
form of official annual reports for certain hospitals. Hence I might
just use available data from (the more unreliable) Wikipedia to estimate
this variable; if not I will omit from presenting this data since it is
extremely incomplete.

------------------------------------------------------------------------

## Diary Entry (Week 10)

### (1) What is the question that you are going to answer? (Answer: One sentence that ends with a question mark that could act like the title of your data story)

What contributes to the difference in wait times for admission to wards
in local public hospitals?

### (2) Why is this an important question? (Answer: 3 sentences, each of which has some evidence, e.g., "According to the United Nations..." to justify why the question you have chosen is important).

The CNA had reported in April 2023 that the median waiting time for ward
admissions had increased from about 5 hours to 7.2 hours during that
period^[1](https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136)^.
The health ministry had explained that it was due to the rise in number
of older patients with more complex medical needs requiring longer
hospitalisation^[1](https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136)^.
Due to the ageing population in Singapore, long waiting time for
admission to wards has been a long-standing issue over the years, and it
was further exacerbated by the Covid-19
pandemic^[2](https://www.channelnewsasia.com/singapore/hospitals-beds-waiting-time-50-hours-admission-ng-teng-fong-sengkang-3014596),[3](https://www.straitstimes.com/singapore/health/most-public-hospitals-saw-more-ae-patients-in-past-week-compared-with-past-month)^.

### (3) Which rows and columns of the dataset will be used to answer this question? (Answer: Actual names of the variables in the dataset that you plan to use)

The main dataset will be the wait times in each hospital across
different dates. Supplementary dataset includes the bed occupancy rate
in each hospital across different dates.

```{r, eval=TRUE,echo=TRUE}
library(tidyverse)
wait_times <- read_csv("../Proj dataset/wt-for-admission-to-ward_week39y2023.csv")
glimpse(wait_times)

bed_occ <- read_csv("../Proj dataset/bed-occupancy-rate_week39y2023.csv")
glimpse(bed_occ)
```

### (4) Challenges and errors that I faced and how I overcame them.

I faced some errors when I tried to read my dataset in R. It was
originally in ".xlsx" format, which resulted in a tibble of 1 x 1 and
unable to display any data. After saving my excel sheet in ".csv"
format, I could read my data but its display was erroneous, because
there were unnecessary table heading and spaces which disrupted the data
structure crucial for a .csv file. After tidying the file by removing
unnecessary portions of the table, I managed to produce the expected
dataset with correct number of columns and rows.

------------------------------------------------------------------------

## Diary Entry (Week 9)

### (1) What is the topic that you have finalized? (Answer in 1 or 2 sentences)

Due to the ageing population, the healthcare sector in Singapore is
facing various challenges, i.e. bed crunch. To gain a deeper insight on
this specific issue, we analyse datasets made available by the Ministry
of Health.

### (2) What are the data sources that you have curated so far? (Answer 1 or 2 sentences).

Two data sources have been curated. <br>

a)  Beds Occupancy Rate.
    **([Link](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/beds-occupancy-rate-(bor)))**
    <br>

b)  Waiting Time for Admission to Ward.
    **([Link](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/waiting-time-for-admission-to-ward))**
