---
author: "Lew Bo Cong"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  df_print: paged
  html_document: null
title: "Challenge-7"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# II. Code to edit and execute using the Code-along-7.Rmd file 
<br>

## A. All about ggplot2 package
<br>

### 1. The Palmer Penguins (Slide #6)

```{r, eval=TRUE,echo=TRUE}
library(tidyverse)
library(palmerpenguins)
glimpse(penguins)
```
<br>

### 2. Palmer Penguins: Plot recreation (Slide #8-18)

```{r, eval=TRUE,echo=TRUE}
# Step 1
ggplot(data = penguins)
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 2
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm))
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 3
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm))
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 4
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm)) +
  geom_point()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 5
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 6
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 7
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 8
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)",
       y = "Bill length (mm)")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 9
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)",
       y = "Bill length (mm)",
       colour = "Species")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 10
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)",
       y = "Bill length (mm)",
       colour = "Species",
       caption = "Source: Palmer Station LTER")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Step 11
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species)) +
  geom_point() +
  labs(title = "Bill depth and length",
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)",
       y = "Bill length (mm)",
       colour = "Species",
       caption = "Source: Palmer Station LTER",
       scale_colour_viridis_d())
```
<br>

### 3. Palmer Penguins: Argument names (Slide #20)

```{r, eval=TRUE,echo=TRUE}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           colour = species)) +
  geom_point() +
  scale_colour_viridis_d()
```
<br>
```{r, eval=TRUE,echo=TRUE}
ggplot(penguins) + # Data layer
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           colour = species) + # Aesthetics layer
              geom_point() + # Geometric layer
                scale_colour_viridis_d()
```
<br>

### 4. Aesthetics options (Slide #22-26)

```{r, eval=TRUE,echo=TRUE}
# Colour
ggplot(penguins) + aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       colour = species) +
  geom_point() + scale_colour_viridis_d()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Shape 1
ggplot(penguins) + aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       colour = species,
                       shape = island) +
  geom_point() + scale_colour_viridis_d()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Shape 2
ggplot(penguins) + aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       colour = species,
                       shape = species) +
  geom_point() + scale_colour_viridis_d()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Size
ggplot(penguins) + aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       colour = species,
                       shape = species,
                       size = body_mass_g) +
  geom_point() + scale_colour_viridis_d()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Alpha
ggplot(penguins) + aes(x = bill_depth_mm,
                       y = bill_length_mm,
                       colour = species,
                       shape = species,
                       size = body_mass_g,
                       alpha = flipper_length_mm) +
  geom_point() + scale_colour_viridis_d()
```
<br>

### 5. Mapping vs Setting (Slide #28)

```{r, eval=TRUE,echo=TRUE}
# Mapping
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm,
      size = body_mass_g,
      alpha = flipper_length_mm) +
  geom_point()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Setting
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point(size = 2, alpha = 0.5)
```
<br>

### 6. Faceting (Slide #29-36)

```{r, eval=TRUE,echo=TRUE}
# Facet 1
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point() +
  facet_grid(species ~ island) # species along the rows, island along the columns
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet 2
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point() +
  facet_grid(species ~ sex) # NA means certain obs with gender not tabulated
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet 3
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point() +
  facet_grid(sex ~ species)
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet 4
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point() +
  facet_wrap(~ species) # only sort by 1 variable
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet 5
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point() +
  facet_wrap(~ species, ncol = 2)
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet 6
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm) +
  geom_point() +
  facet_grid(. ~ species) # grid needs variable for row & column, 
                          # use '.' if don't want to specify one of them
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet & Colour
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm,
      colour = species) +
  geom_point() +
  facet_grid(species ~ sex) +
  scale_colour_viridis_d()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet & Colour, no legend
ggplot(penguins) +
  aes(x = bill_depth_mm,
      y = bill_length_mm,
      colour = species) +
  geom_point() +
  facet_grid(species ~ sex) +
  scale_color_viridis_d() +
  guides(color = "none")
```
<br>

## B. Visualising numeric variables

### 7. The Lending Club (Slides #39-40)

```{r, eval=TRUE,echo=TRUE}
library(openintro)
glimpse(loans_full_schema)
```
<br>
```{r, eval=TRUE,echo=TRUE}
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade,
         state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```
<br>

### 8a. Histogram (Slides #46)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = loan_amount) + 
  geom_histogram() # to know how many times a certain value appears in a variable
                   # or in a distribution (frequency of value)
```
<br>

### 8b. Histogram - varying binwidths (Slides #47-49)

```{r, eval=TRUE,echo=TRUE}
# binwidth = 1000
ggplot(loans) + aes(x = loan_amount) + 
  geom_histogram(binwidth = 1000)
```
<br>
```{r, eval=TRUE,echo=TRUE}
# binwidth = 5000
ggplot(loans) + aes(x = loan_amount) + 
  geom_histogram(binwidth = 5000)
```
<br>
```{r, eval=TRUE,echo=TRUE}
# binwidth = 20000
ggplot(loans) + aes(x = loan_amount) + 
  geom_histogram(binwidth = 20000)
```
<br>

### 8c. Histogram - customisations (Slides #50-52)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) +
  aes(x = loan_amount) + 
    geom_histogram(binwidth = 5000) +
      labs(x = "Loan amount ($)",
           y = "Frequency",
           title = "Amounts of Lending Club loans")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Fill with categorical variable
ggplot(loans) +
  aes(x = loan_amount,
      fill = homeownership) + 
    geom_histogram(binwidth = 5000,
                   alpha = 0.5) +
      labs(x = "Loan amount ($)",
           y = "Frequency",
           title = "Amounts of Lending Club loans")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Facet with categorical variable
ggplot(loans) +
  aes(x = loan_amount,
      fill = homeownership) + 
    geom_histogram(binwidth = 5000) +
      labs(x = "Loan amount ($)",
           y = "Frequency",
           title = "Amounts of Lending Club loans") +
        facet_wrap(~ homeownership, nrow = 3)
```
<br>

### 9a. Density plot (Slides #53)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = loan_amount) + 
  geom_density() # just like histograms but represented by smooth curves instead;
                 # gives probability density of data in each range instead of count
                 # (no. of times values in certain range occur over
                 # total no. of values in the variable)
```
<br>

### 9b. Density plot - varying bandwidths (Slides #54-56)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = loan_amount) + 
  geom_density(adjust = 0.5)
```
<br>
```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = loan_amount) + 
  geom_density(adjust = 1) # default bandwidth
```
<br>
```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = loan_amount) + 
  geom_density(adjust = 2)
```
<br>

### 9c. Density plot - customisations (Slides #57-58)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = loan_amount) + 
  geom_density(adjust = 2) +
      labs(x = "Loan amount ($)",
           y = "Density",
           title = "Amounts of Lending Club loans")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# Add categorical variable
ggplot(loans) + aes(x = loan_amount,
                    fill = homeownership) + 
  geom_density(adjust = 2,
               alpha = 0.5) +
      labs(x = "Loan amount ($)",
           y = "Density",
           title = "Amounts of Lending Club loans",
           fill = "Homeownership")
```
<br>

### 10a. Box plot (Slides #59-60)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = interest_rate) + 
  geom_boxplot() # vertical line in the middle = median, boundaries = IQR,
                 # dots outside = outliers, tips of horizontal line = min & max
```
<br>
```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = annual_income) + 
  geom_boxplot()
```
<br>

### 10b. Box plot - customisations (Slides #61-62)
```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = interest_rate) + 
  geom_boxplot() +
  labs(x = "Interest rate (%)",
       y = NULL,
       title = "Interest rates of Lending Club loans") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```
<br>
```{r, eval=TRUE,echo=TRUE}
# adding categoric variable
ggplot(loans) + aes(x = interest_rate,
                    y = grade) + 
  geom_boxplot() +
  labs(x = "Interest rate (%)",
       y = "Grade",
       title = "Interest rates of Lending Club loans",
       subtitle = "by grade of loan")
```
<br>

### 11. Scatterplot (Slides #63)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = debt_to_income,
                    y = interest_rate) +
  geom_point() # but too concentrated on the left
```
<br>

### 12. Hex plot (Slides #64-65)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = debt_to_income,
                    y = interest_rate) +
  geom_hex() # similar to histogram; 
```
<br>
```{r, eval=TRUE,echo=TRUE}
ggplot(loans %>% filter(debt_to_income < 100)) + aes(x = debt_to_income,
                    y = interest_rate) +
  geom_hex() 
```
<br>

## C. Visualising categoric variables

### 13. Bar plot (Slides #67-71)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = homeownership) +
  geom_bar()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# segmented bar plot
ggplot(loans) + aes(x = homeownership,
                    fill = grade) +
  geom_bar()
```
<br>
```{r, eval=TRUE,echo=TRUE}
# segmented bar plot
ggplot(loans) + aes(x = homeownership,
                    fill = grade) +
  geom_bar(position = "fill")
```
<br>
```{r, eval=TRUE,echo=TRUE}
# customisation of bar plot
ggplot(loans) + aes(y = homeownership,
                    fill = grade) +
  geom_bar(position = "fill") +
  labs(x = "Proportion",
       y = "Homeownership",
       fill = "Grade",
       title = "Grades of Lending Club loans",
       subtitle = "and homeownership of lendee")
```
<br>

## D. Visualising variables of varied types

### 14. Violin plot (Slides #73)

```{r, eval=TRUE,echo=TRUE}
ggplot(loans) + aes(x = homeownership,
                    y = loan_amount) +
  geom_violin()
```
<br>

### 15. Ridge plot (Slides #74)

```{r, eval=TRUE,echo=TRUE}
library(ggridges)
ggplot(loans) + aes(x = loan_amount,
                    y = grade,
                    fill = grade,
                    color = grade) +
  geom_density_ridges(alpha = 0.5)
```
<br>