---
title: "Challenge-3"
author: "Lew Bo Cong"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}

</style>

## I. Questions

#### Question 1: Emoji Expressions

Imagine you're analyzing social media posts for sentiment analysis. If you were to create a variable named "postSentiment" to store the sentiment of a post using emojis ( for positive, for neutral, for negative), what data type would you assign to this variable? Why? (_narrative type question, no code required_)

**Solution:**  I would assign the 'character' data type to this variable, because this is a categorical variable that is ordinal. 


#### Question 2: Hashtag Havoc

In a study on trending hashtags, you want to store the list of hashtags associated with a post. What data type would you choose for the variable "postHashtags"? How might this data type help you analyze and categorize the hashtags later? (_narrative type question, no code required_)

**Solution:**  I would choose the 'character string' data type since it is a nominal variable. I will be able to store this data type in a vector and insert a counting function which allows me to see which hastags are trending.


#### Question 3: Time Traveler's Log
  
You're examining the timing of user interactions on a website. Would you use a numeric or non-numeric data type to represent the timestamp of each interaction? Explain your choice (_narrative type question, no code required_)

**Solution:**  I would use a non-numeric data type, because the timestamp is used as a display and no calculations will be involved.


#### Question 4: Event Elegance

You're managing an event database that includes the date and time of each session. What data type(s) would you use to represent the session date and time? (_narrative type question, no code required_)

**Solution:**  I would use the 'character string' data type to represent both the session date and time.

#### Question 5: Nominee Nominations
You're analyzing nominations for an online award. Each participant can nominate multiple candidates. What data type would be suitable for storing the list of nominated candidates for each participant?  (_narrative type question, no code required_)

**Solution:**  The 'character string' data type will be suitable. 

#### Question 6: Communication Channels
In a survey about preferred communication channels, respondents choose from options like "email," "phone," or "social media." What data type would you assign to the variable "preferredChannel"? (_narrative type question, no code required_)

**Solution:**  I would assign the 'character string' data type.

#### Question 7: Colorful Commentary

In a design feedback survey, participants are asked to describe their feelings about a website using color names (e.g., "warm red," "cool blue"). What data type would you choose for the variable "feedbackColor"? (_narrative type question, no code required_)

**Solution:**  I would choose the 'character string' data type.

#### Question 8: Variable Exploration

Imagine you're conducting a study on social media usage. Identify three variables related to this study, and specify their data types in R. Classify each variable as either numeric or non-numeric.

**Solution:**  The possible three variables are: brand of social media accessed, access count in a day, and total usage time in a day. Their data and variable types are: non-numeric character string, numeric integer, and numeric double, respectively.

#### Question 9: Vector Variety

Create a numeric vector named "ages" containing the ages of five people: 25, 30, 22, 28, and 33. Print the vector.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
ages <- c(25,30,22,28,33)
print(ages)
```

#### Question 10: List Logic

Construct a list named "student_info" that contains the following elements:

  - A character vector of student names: "Alice," "Bob," "Catherine"

  - A numeric vector of their respective scores: 85, 92, 78

  - A logical vector indicating if they passed the exam: TRUE, TRUE, FALSE

Print the list.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
student_info <- list(names=c("Alice","Bob","Catherine"), scores=c(85,92,78), passed=c(TRUE,TRUE,FALSE))
print(student_info)
```

#### Question 11: Type Tracking

You have a vector "data" containing the values 10, 15.5, "20", and TRUE. Determine the data types of each element using the typeof() function.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
data <- c(10,15.5,"20", TRUE)
typeof(data[1])
typeof(data[2])
typeof(data[3])
typeof(data[4])
```

#### Question 12: Coercion Chronicles

You have a numeric vector "prices" with values 20.5, 15, and "25". Use explicit coercion to convert the last element to a numeric data type. Print the updated vector.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
prices <- c(20.5,15,"25")
prices <- as.numeric(prices)
print(prices)
```

#### Question 13: Implicit Intuition

Combine the numeric vector c(5, 10, 15) with the character vector c("apple", "banana", "cherry"). What happens to the data types of the combined vector? Explain the concept of implicit coercion.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
num_vector <- c(5,10,15)
typeof(num_vector)
char_vector <- c("apple","banana","cherry")
typeof(char_vector)
combi_vector <- c(num_vector, char_vector)
typeof(combi_vector)
```

The data types of the combined vector becomes 'character'. Implicit coercion is the automatic conversion of data types by R based on the contents of the vector.

#### Question 14: Coercion Challenges

You have a vector "numbers" with values 7, 12.5, and "15.7". Calculate the sum of these numbers. Will R automatically handle the data type conversion? If not, how would you handle it?

**Solution:** 

```{r, eval=FALSE,echo=TRUE}
# Enter code here
num_vec <- c(7,12.5,"15.7")
sum(num_vec)
```

No, R will not automatically handle the data type conversion. I would do a explicit coercion as follows:
```{r, eval=TRUE,echo=TRUE}
num_vec <- c(7,12.5,"15.7")
num_vec <- as.numeric(num_vec)
sum(num_vec)
```

#### Question 15: Coercion Consequences

Suppose you want to calculate the average of a vector "grades" with values 85, 90.5, and "75.2". If you directly calculate the mean using the mean() function, what result do you expect? How might you ensure accurate calculation?

**Solution:** 

I would expect the result to return 'NA', because the mean() function requires the "grades" vector to be in numeric data type.
```{r, eval=FALSE,echo=TRUE}
# Enter code here
grades <- c(85,90.5,"75.2")
mean(grades)
```

To ensure accurate calculation, I would do an explicit coercion to ensure the data type is numeric.
```{r, eval=TRUE,echo=TRUE}
# Enter code here
grades <- c(85,90.5,"75.2")
grades <- as.numeric(grades)
mean(grades)
```

#### Question 16: Data Diversity in Lists

Create a list named "mixed_data" with the following components:

  - A numeric vector: 10, 20, 30
  
  - A character vector: "red", "green", "blue"

  - A logical vector: TRUE, FALSE, TRUE
  
Calculate the mean of the numeric vector within the list.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
mixed_data <- list(num_vec2=c(10,20,30), char_vec2=c("red","green","blue"), log_vec=c(TRUE,FALSE,TRUE))
mean(mixed_data$num_vec2)
```

#### Question 17: List Logic Follow-up

Using the "student_info" list from Question 10, extract and print the score of the student named "Bob."

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
student_info[["names"]]
student_info[["scores"]][2]
```

#### Question 18: Dynamic Access

Create a numeric vector values with random values. Write R code to dynamically access and print the last element of the vector, regardless of its length.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
rand_vec <- c(2,5,3,15,63,16)
rand_vec[length(rand_vec)]
```

#### Question 19: Multiple Matches

You have a character vector words <- c("apple", "banana", "cherry", "apple"). Write R code to find and print the indices of all occurrences of the word "apple."

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
char_vec2 <- c("apple","banana","cherry","apple")
output_vec <- 1:length(char_vec2)
print(output_vec[char_vec2=="apple"])
```

#### Question 20: Conditional Capture

Assume you have a vector ages containing the ages of individuals. Write R code to extract and print the ages of individuals who are older than 30.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
ages <- c(5,61,72,3,8,4,68,56,4,34,12,37)
ages[ages>30]
```

#### Question 21: Extract Every Nth

Given a numeric vector sequence <- 1:20, write R code to extract and print every third element of the vector.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
num_seq <- 1:20
num_seq[seq(from=3,to=20,by=3)]
```

#### Question 22: Range Retrieval

Create a numeric vector numbers with values from 1 to 10. Write R code to extract and print the values between the fourth and eighth elements.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
num_seq2 <- 1:10
num_seq2[4:8]
```

#### Question 23: Missing Matters

Suppose you have a numeric vector data <- c(10, NA, 15, 20). Write R code to check if the second element of the vector is missing (NA).

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
num_vec2 <- c(10, NA, 15, 20)
print(is.na(num_vec2[2]))
```

#### Question 24: Temperature Extremes

Assume you have a numeric vector temperatures with daily temperatures. Create a logical vector hot_days that flags days with temperatures above 90 degrees Fahrenheit. Print the total number of hot days.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
temp_vec <- c(60,20,75,86,90,54,95,91,79)
hot_days <- temp_vec[1:length(temp_vec)]>90
sum(hot_days)
```

#### Question 25: String Selection

Given a character vector fruits containing fruit names, create a logical vector long_names that identifies fruits with names longer than 6 characters. Print the long fruit names.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
fruits_vec <- c("apple","banana","pineapple","grapes","kiwi")
long_names <- nchar(fruits_vec[1:length(fruits_vec)])>6
print(fruits_vec[long_names])
```

#### Question 26: Data Divisibility

Given a numeric vector numbers, create a logical vector divisible_by_5 to indicate numbers that are divisible by 5. Print the numbers that satisfy this condition.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
num_vec3 <- c(3,5,2,50,25,34,60,73,10,15)
divisible_by_5 <- num_vec3%%5==0
print(num_vec3[divisible_by_5])
```

#### Question 27: Bigger or Smaller?

You have two numeric vectors vector1 and vector2. Create a logical vector comparison to indicate whether each element in vector1 is greater than the corresponding element in vector2. Print the comparison results.

**Solution:** 

```{r, eval=TRUE,echo=TRUE}
# Enter code here
vector1 <- c(2,3,51,2,4,7,5,2)
vector2 <- c(3,15,6,4,0,5,6,2)
greater_vec <- vector1 > vector2
print(greater_vec)
```
