---
title: "Diary Entry"
author: "Lew Bo Cong"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Submission (Week 13)

#### (1) What is the theme of your data story?

In the 21st century, the human life very often begins (birth) and ends
(death) at the hospital, especially if we are fortunate enough to live
in a developed city. Once born in Singapore, we receive the health
booklet that accompanies us throughout our childhood and adolescence
period. The technological revolution in recent years had created the
National Electronic Health Record that now stores our healthcare history
(*NEHR*, n.d.). And to mitigate the ever-increasing inflation and
support the necessary and eventual healthcare expenditure, the
government had also established MediSave (*MOH \| MediSave,* n.d.).
These points show that healthcare and its related services, institutions
and policies are deeply integrated in our lives, regardless of our
health status. Consequently, the theme of my data story revolves around
the healthcare industry in Singapore, focusing on the problem of long
waiting time for admission to a ward in our public hospitals.

#### (2) Why is it important to address this question?

The CNA reported that the median waiting time for ward admissions had
increased from about 5 hours to 7.2 hours in April 2023 (CNA, 2023). The
health ministry explained that it was due to more older patients with
increasingly complex medical needs requiring longer hospitalisation
(CNA, 2023). Due to the ageing population in Singapore, long waiting
time for admission to wards have been a long-standing issue over the
years, which was further exacerbated by the Covid-19 pandemic (Lim,
2022; Ng, 2021). In our fast-paced lifestyle, we are accustomed to
receiving instant services, like fast food, news from social media, and
punctual public transportation, hence we turn impatient and anxious when
forced to experience delays. These experiences worsen when we, or our
loved ones, are in pain while waiting for treatment or a ward. As
interviewed by the CNA (2022), Mr Ng was "very worried" when his mother
with advanced dementia had to wait 25 hours alone for a hospital bed.
Experiencing an unexpected medical condition can be a stressful
experience, and needing to wait long durations for a ward adds on to the
distress.

#### (3) Why do you think the data sources that you have curated can help you answer the question?

The Ministry of Health collates raw data on the waiting time for
admission to ward in the public hospitals in Singapore, made available
publicly
[here](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/waiting-time-for-admission-to-ward).
The dataset consists of daily waiting times at the 50th percentile in
2023, and data was manipulated to identify the peak periods and hospital
location(s) with consistently high waiting times. Additional data sets
such as the [Bed Occupancy
Rate](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/beds-occupancy-rate-(bor))
and [Attendances at Emergency Medicine
Departments](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/attendances-at-emergency-medicine-departments)
(EMD) -- a total of 3 variables -- were curated to seek possible
correlations that explain the long waiting times faced in the hospitals.
These datasets were strategically curated to test the hypothesis that
long waiting time is contributed by "higher bed utilisation" and the
phenomenon of patients visiting the emergency department for non-urgent
conditions (CNA, 2023).

#### (4) What are the insights from the data and how are they depicted in plots?

All datasets were manipulated and depicted in suitable plots in the
shiny app. Firstly, general comparisons in the 3 variables across the
different hospitals were made into line charts to illustrate absolute
data and box plots to visualise descriptive statistics such as the
median, interquartile range and number of outliers. Because everyone
experiences a unique set of conditions and requires different forms of
treatment, the usage pattern of hospital facilities will fluctuate
across different locations and time periods. Taking the waiting time
boxplots as an example, we can see that Khoo Teck Puat Hospital (KTPH)
had the biggest interquartile range in February, but in April Ng Teng
Fong General Hospital (NTFGH) had the biggest interquartile range
instead, both exceeding the 10-hour mark. However, it appears that Tan
Tock Seng Hospital (TTSH) had a relatively small interquartile range
that is under 10 hours. This might mean that the standard operating
procedures TTSH implement could produce a reliable system with lower
waiting times.

Secondly, we focus on the individual hospital statistics, with the help
of a calendar heatmap that allows us to spot the longest or shortest
variable, or patterns across the year at a glance. For example, KTPH
experienced the longest waiting time between February to May, with
waiting times peaking in March; EMD attendance tend to spike every
Monday across all the hospitals. To detect correlations between these 3
variables, illustrative data visualisation was done by plotting the
waiting time line chart over bed occupancy rate or EMD attendance bar
chart, and a scatterplot of the respective two variables with the
correlation coefficient calculated and displayed. The correlation
coefficient for each month and each hospital was calculated and
displayed in a table in the website for easier visualisation and
analysis. From this table, we can observe fluctuations in the
correlation coefficients too, but it appears that Singapore General
Hospital (SGH) had the highest maximum and minimum correlation
coefficient of 0.8 and 0.6 respectively for the correlation between
waiting time and bed occupancy rate, as compared to the other hospitals.
This might postulate that bed occupancy rate is a good predictor of
waiting time for admission to a ward, specifically in SGH.

#### (5) How did you implement this entire project? Were there any new concepts that you learnt to implement some aspects of it?

This entire project was implemented using the following framework: be
inspired --\> experiment --\> debug --\> google for help --\> attempt to
fix --\> celebrate or change tactics. With the help of Google, I visited
other websites (either made with Quarto or not) to seek inspiration for
my website structure. Most importantly, I also came across a
[blog](https://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
illustrating popular ways of data visualisation, forming the basis of
how I wanted to manipulate and present my raw data in my shiny app. The
Drag and Drop shiny dashboard introduced by the teaching team was
helpful in kickstarting my shiny app development, but the resulting code
created became too messy, so I stopped using the dashboard and manually
wrote the code for my shiny app instead. By analysing the code written
by the dashboard program, I had a better understanding of the shiny app
structure -- i.e. the variable IDs in the UI that my server required to
produce the respective plots -- which allowed me to experiment with all
the different plots that I wanted to present. The problem of messy and
extensive codes persisted, which motivated me to leverage the power of
functions for efficient codes, as seen in my app.R code script. An
additional example is replacing the long "if-else" statements with the
simple "switch()" function to replace variables as desired. However, a
big problem that I faced was trying to personalise the website I
envisioned, mainly because I do not have the background in customising
bootstrap and css. I tried to learn from the W3Schools tutorials, but I
find the learning curve quite steep within such a small timeframe. At
least, I could insert a cover image with overlaying text using css,
which is something that I visualised. Therefore, I have to change
tactics and tap on the built-in website structure already available in
Quarto. Nonetheless, I managed to implement new concepts such as
additional functions in tidying data, including "gather()",
"separate()", "cbind()"; utilising factors to transform data and create
a calendar heatmap; additional arguments to customise ggplot and add
hover tooltips; combining linechart with barchart; creating scatterplots
including coefficient correlation calculation. These new concepts
complement the linecharts, boxplots and tidy-data manipulation learnt in
class, thereby presenting my data story that was not only interactive
but meaningful.

(1195 words)

### References

CNA. (2023, April 25). Median wait time for admission to hospital wards
has gone up to 7.2 hours: MOH. *CNA*.
<https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136>

Lim, V. (2022, October 20). Longer waiting times at hospitals with some
patients told to wait up to 50 hours for a bed. *CNA*.
<https://www.channelnewsasia.com/singapore/hospitals-beds-waiting-time-50-hours-admission-ng-teng-fong-sengkang-3014596>

*MOH \| MediSave*. (n.d.). Retrieved 24 November 2023, from
<https://www.moh.gov.sg/healthcare-schemes-subsidies/medisave>

*NEHR*. (n.d.). Retrieved 24 November 2023, from
<https://www.synapxe.sg/healthtech/national-programmes/national-electronic-health-record-nehr>

Ng, K. G. (2021, September 20). Longer wait for admission with most
public hospitals seeing more A&E patients, Covid-19 cases. *The Straits
Times*.
<https://www.straitstimes.com/singapore/health/most-public-hospitals-saw-more-ae-patients-in-past-week-compared-with-past-month>

------------------------------------------------------------------------

## Diary Entry (Week 12)

### (1) Challenges and errors that I faced and how I overcame them.

1)  My Shiny app couldn't be published on shinyapps.io. I tried to
    search online based on the error "Application failed to start, Exit
    Code 1" and read that it was due to too much memory being used. I
    assumed this was the error and searched the internet on how to fix
    this, and learnt that I could use the googlesheets4 package to
    create, read, and write data tables and store them remotely on
    Google Drive (which is something I have learnt outside of class).
    Turns out, this was not the reason for the error and thanks to prof
    at the consultation session, the issue was fixed just by placing my
    data .csv files in the same folder as my app.R file.

2)  My code was too long and messy, because the Drag & Drop Shiny
    Dashboard writes the code in a linear manner that appends the code
    chunk depending on my actions in the Dashboard. Hence, I searched
    online for tips to structure the Shiny app and learnt that we can
    create functions for repetitive UI and Server codes.

-   I identified repetitive sections like sliderInput, selectInput,
    checkboxGroupInput, plotlyOutput, nav_panel (mainly in the UI
    section) and created functions with appropriate arguments so that I
    can define them strategically and automatically create the unique
    variable and ID names for downstream usage. I used the switch()
    function (which is something I have learnt outside of class) instead
    of the if() function to make my code clearer and straightforward.
-   I faced a problem when I created functions to be used in the Server,
    mainly to help me plot the required linecharts and boxplots for each
    data groups (i.e. Waiting Time, Bed Occupancy Rate, EMD Attendance)
    that I have. As the data required for each plot and data groups
    differs, I tried to define a common variable (i.e. 'datasource')
    that can be given a unique variable name (i.e. 'wait_duration',
    'occ_rate_num', 'attendance' --\> column variables from my data
    table) using the switch() function, but the error: \<object '' not
    found\> keeps appearing. I realised it is because the object, which
    is the common variable (i.e. 'datasource'), has not been defined so
    I was unable to call it. My solution was to directly apply the
    switch() function at where the common variable should be placed,
    instead of trying to define it in the beginning of the code.

------------------------------------------------------------------------

## Diary Entry (Week 11)

### (1) List the visualizations that you are going to use in your project (Answer: What are the variables that you are going to plot? How will it answer your larger question?)

#### **For multiple hospitals**

-   Line charts
    -   plot Dates against Waiting time, Bed Occupancy Rate, and EMD
        Attendance number, in varying weeks or months.
-   Box plots (Horizontal)
    -   plot Hospitals against Waiting time, Bed Occupancy Rate, and EMD
        Attendance number, in varying weeks, months, or years.
-   Bar charts (Horizontal)
    -   plot Hospitals against Bed Occupancy Rate, and EMD Attendance
        number, on different days.

#### **For individual hospital**

-   Calendar heatmap using facet wrap
    -   plot Waiting time arranged by Days and Week number.
-   Line chart overlaying bar chart
    -   plot Dates against Waiting time (line) and Bed Occupancy Rate
        (bar), in varying weeks or months.
    -   plot Dates against Waiting time (line) and EMD Attendance number
        (bar), in varying weeks or months.
-   Scatterplot
    -   plot Waiting time against Bed Occupancy Rate, in varying weeks,
        months or year.
    -   plot Waiting time against EMD Attendance number, in varying
        weeks, months or year.

By utilising the above-mentioned data visualisations, we can look for
possible correlations and patterns that explain the long waiting times
for admission to a ward.

### (2) How do you plan to make it interactive? (Answer: features of ggplot2/shiny/markdown do you plan to use to make the story interactive)

I plan to use shiny to create an interactive app that allows the user to
input a date range and select hospitals to be drawn on the
above-mentioned plots. I wish to use ggplot2 to display relevant
information on the screen while the user hovers over the data points.

### (3) What concepts incorporated in your project were taught in the course and which ones were self-learnt? (Answer: Create a table with topics in one column and Weeks in the other to indicate which concept taught in which week is being used. Leave the entry of the Week column empty for self-learnt concepts)

+--------------------------------------------+---------------+
| Topics                                     | Weeks         |
+============================================+===============+
| -   Data variables and their types:        | 3             |
|     transforming data into correct type    |               |
|     for manipulation                       |               |
|                                            |               |
|     -   e.g. using as.numeric()            |               |
+--------------------------------------------+---------------+
| -   Using variable type `factor` and       | self-learnt   |
|     `date` to help order categorical       |               |
|     variables                              |               |
|                                            |               |
|     -   e.g. fct_inorder( )                |               |
|                                            |               |
|     -   fct_reorder()                      |               |
|                                            |               |
|     -   parse_date_time()                  |               |
|                                            |               |
|     -   %U                                 |               |
+--------------------------------------------+---------------+
| -   Manipulating Data: understood what is  | 4, 9          |
|     exactly a tidy dataset, and the        |               |
|     operations to tidy, manipulate and     |               |
|     reshape data, especially via pipe      |               |
|     operator for cleaner code              |               |
|                                            |               |
|     -   e.g. mutate()                      |               |
+--------------------------------------------+---------------+
| -   Using other functions to tidy data     | self-learnt   |
|                                            |               |
|     -   e.g. gather()                      |               |
|                                            |               |
|     -   relocate()                         |               |
|                                            |               |
|     -   cbind()                            |               |
+--------------------------------------------+---------------+
| -   Data visualisation via ggplot2:        | 7             |
|     applying the different ways to plot    |               |
|     data for visualisation                 |               |
+--------------------------------------------+---------------+
| -   Learnt and used additional methods of  | self-learnt   |
|     visualising data                       |               |
|                                            |               |
|     -   e.g. various visualisations        |               |
|                                            |               |
|     -   using facetwrap to create a        |               |
|         calendar                           |               |
|                                            |               |
|     -   pie chart                          |               |
|                                            |               |
|     -   overlay line chart over boxplot    |               |
+--------------------------------------------+---------------+

### (4) Challenges and errors that I faced and how I overcame them.

The main challenges and errors were trying to present data in a logical
and meaningful way, by customising conventional data plots in the way I
want. Whenever I faced the issue of needing to add or remove features in
my plots, I will search for the solution on Google and experiment it in
my code. <br> I would like to show the bed capacity of each public
hospital, perhaps on a map, but this data is only available reliably and
accurately in the form of official annual reports for certain hospitals.
Hence I might just use available data from (the more unreliable)
Wikipedia to estimate this variable; if not I will omit from presenting
this data since it is extremely incomplete.

------------------------------------------------------------------------

## Diary Entry (Week 10)

### (1) What is the question that you are going to answer? (Answer: One sentence that ends with a question mark that could act like the title of your data story)

What contributes to the difference in wait times for admission to wards
in local public hospitals?

### (2) Why is this an important question? (Answer: 3 sentences, each of which has some evidence, e.g., "According to the United Nations..." to justify why the question you have chosen is important).

The CNA had reported in April 2023 that the median waiting time for ward
admissions had increased from about 5 hours to 7.2 hours during that
period^[1](https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136)^.
The health ministry had explained that it was due to the rise in number
of older patients with more complex medical needs requiring longer
hospitalisation^[1](https://www.channelnewsasia.com/singapore/moh-hospital-waiting-times-7-hours-covid-19-3442136)^.
Due to the ageing population in Singapore, long waiting time for
admission to wards has been a long-standing issue over the years, and it
was further exacerbated by the Covid-19
pandemic^[2](https://www.channelnewsasia.com/singapore/hospitals-beds-waiting-time-50-hours-admission-ng-teng-fong-sengkang-3014596),[3](https://www.straitstimes.com/singapore/health/most-public-hospitals-saw-more-ae-patients-in-past-week-compared-with-past-month)^.

### (3) Which rows and columns of the dataset will be used to answer this question? (Answer: Actual names of the variables in the dataset that you plan to use)

The main dataset will be the wait times in each hospital across
different dates. Supplementary dataset includes the bed occupancy rate
in each hospital across different dates.

```{r, eval=TRUE,echo=TRUE}
library(tidyverse)
wait_times <- read_csv("../Proj dataset/wt-for-admission-to-ward_week39y2023.csv")
glimpse(wait_times)

bed_occ <- read_csv("../Proj dataset/bed-occupancy-rate_week39y2023.csv")
glimpse(bed_occ)
```

### (4) Challenges and errors that I faced and how I overcame them.

I faced some errors when I tried to read my dataset in R. It was
originally in ".xlsx" format, which resulted in a tibble of 1 x 1 and
unable to display any data. After saving my excel sheet in ".csv"
format, I could read my data but its display was erroneous, because
there were unnecessary table heading and spaces which disrupted the data
structure crucial for a .csv file. After tidying the file by removing
unnecessary portions of the table, I managed to produce the expected
dataset with correct number of columns and rows.

------------------------------------------------------------------------

## Diary Entry (Week 9)

### (1) What is the topic that you have finalized? (Answer in 1 or 2 sentences)

Due to the ageing population, the healthcare sector in Singapore is
facing various challenges, i.e. bed crunch. To gain a deeper insight on
this specific issue, we analyse datasets made available by the Ministry
of Health.

### (2) What are the data sources that you have curated so far? (Answer 1 or 2 sentences).

Two data sources have been curated. <br>

a)  Beds Occupancy Rate.
    **([Link](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/beds-occupancy-rate-(bor)))**
    <br>

b)  Waiting Time for Admission to Ward.
    **([Link](https://www.moh.gov.sg/resources-statistics/healthcare-institution-statistics/waiting-time-for-admission-to-ward))**
